{\tt APFELgrid}: a fast interface to hadronic observables for fits of parton densities

### 2016-05-04 paper.tex
section 2 edits

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,,
\end{equation}
where
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross section with the
interpolating polynomials. This information may be stored in a
precomputed look-up table.  The final expression for the cross section
in Eq.~(\ref{eq:applconv}) is therefore a considerably simpler task to perform
inside a fit than the direct evaluation of the double convolution.

@@

A number of tools, $e.g.$ {\tt
  APFEL},
{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay},
are available which perform PDF evolution via an analogous
procedure. {\tt APFEL} in particular, is able to provide the so-called
\textit{evolution operator}. The evolution operator, is computed
assuming that the DGLAP evolution can be encoded in a single operator
whose action on the intial scale PDFs proceduces the final scale
PDFs. Taking $Q_0$ as an initial scale and $Q_\tau$ as a final scale,
on the $x$-space grid one can write:
\begin{equation}\label{eq:fastPDFfinal_recalled}
  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
  \sum_\beta A^\tau_{\alpha\beta, ik}\;
  f_k(x_\beta,Q^2_0)\,, 
\end{equation}
where the operator $A$ is the solution of the DGLAP evolution equation
with baundary condition:
\begin{equation}
A^\tau_{\alpha\beta, ik} =
\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
\end{equation}

Given the evolution operator defined in
Eq.~(\ref{eq:fastPDFfinal_recalled}), we may replace the
(general-scale) PDFs used in the subprocess parton density
Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions like so:
\begin{equation}\label{eq:FK1}
\begin{array}{rcl}
F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
                               \sum_{\delta,\gamma} C^{(s)}_{ij}
                               \left[  A^\tau_{\alpha\delta ik}\;
                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
  \\
\\
&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
\end{array}
\end{equation}
where the object:
\begin{equation}
  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
  A^\tau_{\beta\gamma jl}\,,
\end{equation}
combines the operations of subprocess density construction and PDF
evolution. Going further it is possible to then use the expression for
subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
cross section calculation:
\begin{equation}
\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
\sum_{s} \sum_{\alpha,\beta}
\sum_{\tau} 
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
\end{equation}
Performing some further contractions it is possible to obtain an
extremely compact expression for the calculation of the cross section
in question, in terms of only the initial-scale PDFs and summing only
over the initial scale interpolating $x$-grid and the incoming parton
flavors:
\begin{equation}\label{eq:FK}
  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
\end{equation}
where the object:
\begin{equation}\label{eq:FKTable}
  \widetilde{W}_{kl,\delta\gamma} = \sum_p\sum_{s}\sum_{\alpha,\beta} \sum_{\tau}
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
\end{equation}
is referred to here as an {\tt FK} table, and combines the information
stored in {\tt APPLgrid}-style interpolated weight grids with
analogously interpolated DGLAP evolution operators. This combination
enables for a maximally computationally-efficient expression for the
calculation of observables at hadron colliders under PDF variation.

>>>>

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p \sum_{s}\sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,,
\end{equation}
where
\begin{equation} \label{eq:applgrid}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross section with the
interpolating polynomials. This information may be stored in a
precomputed look-up table.  The final expression for the cross section
in Eq.~(\ref{eq:applconv}) is therefore a considerably simpler task to perform
inside a fit than the direct evaluation of the double convolution.

@@

A number of tools ($e.g.$ {\tt
  APFEL},
{\tt HOPPET}~\cite{Salam:2008qg} and {\tt QCDNUM}~\cite{Botje:2010ay}) 
are available which perform PDF evolution via an analogous
interpolation procedure. In such a way PDFs at a general scale $Q_\tau$ may be
expressed as a product of PDFs at some initial, fitting scale $Q_0$ and an
\textit{evolution operator} obtained by the solution of the DGLAP equation.
\begin{equation}\label{eq:fastPDFfinal_recalled}
  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
  \sum_\beta A^\tau_{\alpha\beta, ik}\;
  f_k(x_\beta,Q^2_0)\,, 
\end{equation}
where latin indices run over PDF flavour, greek indices run over points in an initial-scale interpolating $x$-grid and
the evolution operator $A$ may be accessed directly in the {\tt APFEL} package. Given this operator, we may replace the
(general-scale) PDFs used in the subprocess parton density
Eq.~(\ref{eq:APPLsubproc}) with their equivalent expressions evaluated at the fitting scale as
\begin{equation}\label{eq:FK1}
\begin{array}{rcl}
F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
                               \sum_{\delta,\gamma} C^{(s)}_{ij}
                               \left[  A^\tau_{\alpha\delta ik}\;
                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
  \\
\\
&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
\end{array}
\end{equation}
with the object
\begin{equation}
  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
  A^\tau_{\beta\gamma jl}\,,
\end{equation}
combining the operations of subprocess density construction and PDF
evolution. Going further and using the expression for
subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
cross section calculation we obtain
\begin{equation}
\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
\sum_{s} \sum_{\alpha,\beta}
\sum_{\tau} 
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
\end{equation}
Performing some further contractions it is possible to obtain an
extremely compact expression for the calculation of the cross section
in question, in terms of only the initial-scale PDFs and summing only
over the initial scale interpolating $x$-grid and the incoming parton
flavors:
\begin{equation}\label{eq:FK}
  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
\end{equation}
where the object:
\begin{equation}\label{eq:FKTable}
  \widetilde{W}_{kl,\delta\gamma} = \sum_p\sum_{s}\sum_{\alpha,\beta} \sum_{\tau}
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
\end{equation}
is referred to here as an {\tt FK} table, and combines the information
stored in {\tt APPLgrid}-style interpolated weight grids with
analogously interpolated DGLAP evolution operators. This combination
enables for a maximally efficient expression for the
calculation of observables at hadron colliders under PDF variation.
